---
title: Installation
output:
  bookdown::html_document2:
    theme: flatly
    number_sections: false
pkgdown:
  as_is: true
---

<style>
.contents .page-header {
    display: none;
  } 
</style>

<div class = "article_container">

# Installation
<hr>

**epidemia** requires R v3.5.0 or above. The latest version of R can be downloaded
[here](https://cloud.r-project.org/bin/).

Models are fit with [**Stan**](https://mc-stan.org) through the
[**cmdstanr**](https://mc-stan.org/cmdstanr/) interface. Unlike the previous
**rstan**-based versions of the package, the Stan models are **not** compiled
when **epidemia** is installed. Instead they are compiled the first time they
are used and then cached, so the one-off C++ compilation happens at run time
rather than at install time.

## 1. Install cmdstanr and CmdStan

```{r install-cmdstanr, eval = FALSE}
# install cmdstanr
install.packages(
  "cmdstanr",
  repos = c("https://stan-dev.r-universe.dev", getOption("repos"))
)

# install CmdStan itself (requires a working C++ toolchain: on macOS the
# Xcode command line tools, on Windows RTools, on Linux gcc/g++)
cmdstanr::check_cmdstan_toolchain(fix = TRUE)
cmdstanr::install_cmdstan(cores = parallel::detectCores())
```

See the [cmdstanr getting-started guide](https://mc-stan.org/cmdstanr/articles/cmdstanr.html)
if you run into toolchain issues.

## 2. Install epidemia

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("mlgh-sg/epidemia2")
```

will install the latest development version of **epidemia**.

## 3. (Optional) Precompile the Stan models

The Stan programs are compiled automatically the first time `epim()` is called.
To move that one-off cost to a convenient time (for example straight after
installation, or after upgrading CmdStan), run:

```{r compile, eval = FALSE}
epidemia::compile_epidemia()
```

Vignettes are not currently included in the package because they are
computationally demanding, and are best viewed online.

</div>
